---
title: "General Social Survey"
author: "Pedro Hernandez"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"


```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.

how are the "voted in 2016" and "voted in 2012" related?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
use categorical variables and a chi squared test
the hypothesis that voting in 2016 is/is not independent of voting in 2012

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
what we found was that if someone was willing to vote in 2012, the would probably vote in 2016 as well and if someone was unlikely to vote in 2012 they were likely in 2016
### Descriptive Results
we use barchart to show the voting trends between 2012 and 2016


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent" ,xlab='voted in 2016')
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
mumerical summary is a cross tabulation
```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)

```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```
the hypothesis was that voting in 2012 is related to voting in 2016, we did a chi square and p value is 0 and we reject the null hypothesis and what we found that if you wre likely to vote in 2012 you were also in 2016
# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)
how are believing in the big bang theory and post life related?

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
use categorical variables and a chi squared test
the hypothesis that believing in the big bang theory is independent to the post life

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
NUMERICALLY AND GRAPHICALLY THERE IS A BIG GAP BETWEEN THOSE WHO BELIEVE IN THE BIG BANG THEORY AND THE POSTLIFE SO WE ACCEPT OUR NULL HYPOTHESIS THAT THERE IS NO RELATION
### Descriptive Results
we use bar chart to compare trends in postlife and big bang theory
#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->



```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent" ,xlab='BELIEVE IN POSTLIFE')
```
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
xtabs(~BIGBANG+POSTLIFE, data=mydata3)

```
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
votetab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(votetab)
```
THERE IS NO RELATION BETWEEN BELIEVING IN POSTLIFE AND BIG BANG THEORY THEREFORE WE ACCEPT OUR NULL HYPOTHESIS
# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).
WHAT IS THE RELATIONSHIP BETWEEN BELIEVING IN THE DEATH PENALTY AND WETHER THE COURTS ARE HARSH ENOUGH?
## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
se categorical variables and a chi squared test
the hypothesis that THE DEATH PENALTY IS THE BELIEF THAT THE DEATH PENALTY IS RIGHT IS INDEPENDANT FROM WETHER THE COURTS ARE HARSH ENOUGH
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
IN THIS COMPARISON WE FOUND THAT THOS WHO BELIEVE THAT THE DEATH PENALTY WAS RIGHT, ALSO TENDED TO BELIEVE THAT THE COURTS WERE NOT HARSH ENOUGH AND THOSE WHO DID NOT AGREE WITH THE DEATH PENALTY TENDED TO BELIEVE THAT THE COURTS WERE TOO HARSH
### Descriptive Results
WE USED THE BAR CHART TO COMPARE THE RELATIONSHIP BETWEEN BELIEVING IN THE DEATH PENALTY AND FIGURING IF THE COURTS WERE HARSH OR NOT
#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~CAPPUN+COURTS, data=mydata4, type = "percent" ,xlab='VIEW ON COURTS')
```
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
xtabs(~CAPPUN+COURTS, data=mydata4)

```
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata4)
votetab=xtabs(~CAPPUN+COURTS, data=mydata4)
fisher.test(votetab)
```
WE CAN CONLUDE THAT OUR NULL HYPOTHESIS SHOULD BE IGNORED AS WE CAN SEE THAT THERE IS A RELATIONSHIP BETWEEN THE BELIEFS OF THE HARSHNESS OF THE COURTS AND THE ACCEPTANCE OF THE DEATH PENALTY
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
to conclude, I believe there is some relationships between some of the variables, while the others appear to show the opposite, it all depends on what the variables are and wether they can even have a relationship.
